### 2016-09-26 HRI/Parts/Antecedents.tex
Small correction in the Antecedents, which is a description of the experiment, the system. There is still need work on the case study description.

An experiment was designed to assess precise values for linear velocity, angular velocity, oscillation angle, direction and orientation that could be used to express happiness, angry, fear and sadness, which correspond to four basic emotions suggested by Ekman~\cite{Ekman2001}. These features were selected after the study on emotion projection in robotics, humans and previous case studies. The robotic platform used in the experiment is holonomic, which are characterized by the possibility to move in any direction without the necessity to have a specific orientation, i.e., they are free to move taking any desired orientation. The experiment was performed at Politecnico di Milano, campus Leonardo during the months of June and July of 2015.  A total of 49 volunteers were involved: 12 female and 37 male. The average age of the participants was 25.28 with standard deviation of 2.8, with a minimum age of 20 and maximum of 32.

For each table were calculated the mean, standard deviation, and median. It was not possible to use  ANOVA test over the data because the assumption of normality is not achieved in the collected data. This was check using the Shapiro-Wilk Test.
Additionally, a contingency table for each emotion was generated in each treatment as it is depicted in  Table~\ref{table:table_contingency},  where the intensity for the other emotions is calculated as the mean of them, including the option of ''other''. For all tables, including the contingency, were calculated the Krippendorff's alpha agreement~\cite{Krippendorff2007} ($\alpha$), which is a reliability coefficient to measure the agreement among different participants. Unlike other coefficients (Kappa), $\alpha$ is a generalization of several known reliability indices, and it applies to:

>>>>

An experiment was done to assess precise values for linear velocity, angular velocity, oscillation angle, direction and orientation that could be used to express happiness, angry, fear and sadness, which correspond to four basic emotions suggested by Ekman~\cite{Ekman2001}. These features were selected after study emotion projection in robotics, humans (i.e. previous work and theatre) and previous case studies. The robotic platform used in the experiment is holonomic, which are characterized by the possibility to move in any direction without necessity to have a specific orientation, i.e., they are free to move taking any desired orientation. The experiment was performed at the university campus during the months of June and July of 2015.  A total of 49 volunteers were involved: 12 female and 37 male. The average age of the participants was 25.28 with standard deviation of 2.8, with a minimum age of 20 and maximum of 32.

A Google form was used to collect participants' answers. The form included the four emotions plus two mental states (i.e. excitement and tenderness) and option ''other''. These mental state were added to see if people could confuse the desired emotion with them. Each participant was exposed to twenty over one hundred and ninety five possible treatments. The whole process, including a brief explanation and assessment lasted  from 10 to 15 minutes. This was decided because each subject was a volunteer and would not perceive any monetary remuneration, so the time dedicated to the experiment had to be kept limited. The twenty treatments were selected picking a number without replacement from 1 to 195. For each presented treatment, participants were asked to give an intensity perceived for each option in the form.

After collecting all the data, it was created a table for each treatment. Each table contain the following information mean, standard deviation, and median. ANOVA test was not possible to be used over the data because the assumption of normality is not achieved in the collected data. This was check using the Shapiro-Wilk Test.
Additional to these table, a contingency table for each emotion was generated in each treatment as it is depicted in  Table~\ref{table:table_contingency},  where the intensity for the other emotions is calculated as the mean of them, including the option of ''other''. For all tables, including the contingency, were calculated the Krippendorff's alpha agreement~\cite{Krippendorff2007} ($\alpha$), which is a reliability coefficient to measure the agreement among different participants. Unlike other coefficients (Kappa), $\alpha$ is a generalization of several known reliability indices, and it applies to:

### 2016-09-26 HRI/Parts/CaseStudy.tex
Changes in both articles.
Small correction in the Antecedents, which is a description of the experiment, the system. There is still need work on the case study description.

The case study was done at Researcher's Night 2014 with two main objectives (i) cross-validate the findings obtained from the experiment, and (ii) use the Emotional Enrichment System to verify whether the participants would prefer scenes when the robot expresses emotions or rather moves without any emotion expression.

@@

This case study uses the results obtained in the experiment and was designed to have two parts (i) emotion expression through changes in linear velocity, angular velocity, oscillation angle, orientation and direction. 
%Figure~\ref{fig:example_movements} presents an example of these variables. 
(ii) Presentation of a small scene to verify whether the participants would prefer scenes when the robot shows emotions or not. the Emotional Enrichment System was used in both cases (with and without emotions). Two web-cams and eight Alvar tags were added to use Kalman filter to improve robots localization in the stage. The detection of the AR tags is done through the use of the ROS package ar\_track\_alvar~\cite{artag2015}. The distribution of the web-cams and the tags are depicted in Figure~\ref{fig:setup_fourth}.

@@

\begin{table}
\centering
\small
\caption{Parameters' values selected from the experiment.}
		\label{table:selected_fourth}
		\begin{tabular}{|c|c|c|c|c|c|c|}
			\hline
\rotatebox{90}{\textbf{Emotion } }&
\rotatebox{90}{\textbf{Direction  ($rad$)}}&
\rotatebox{90}{\textbf{Orientation ($rad$)} }&
\rotatebox{90}{\textbf{Linear Velocity ($mm/s$) }}&
\rotatebox{90}{\textbf{Angular Velocity ($rad/s$) }}&
\rotatebox{90}{\textbf{Angle ($rad$)}}&
\rotatebox{90}{\textbf{Treatment ID}}\\	
			\hline
			\multirow{2}{*}{Happiness}&$0$&$0$&$500$&$3$&$0.349$&29\\
			\cline{2-7}
			&$0$&$0$&$900$&$3$&$0.174$&38\\
			\hline
			\multirow{2}{*}{Anger}&$\pi$&$0$&$500$&$3$&$0.087$&144\\
			\cline{2-7}
			&$0$&$0$&$900$&$1$&$0.087$&31\\
			\hline
			\multirow{2}{*}{Fear}&$\pi$&$\pi$&$900$&$2$&$0.174$&113\\
			\cline{2-7}
			&$\pi$&$\pi$&$500$&$2$&$0.087$&102\\
			\hline
			\multirow{2}{*}{Sadness}&$\pi$&$0$&$200$&$1$&$0.349$&130\\
			\cline{2-7}
			&$0$&$\pi$&$200$&$1$&$0.349$&52\\
			\hline
			\end{tabular}
\end{table}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/FourthCaseScene.png} 
	\caption{Stage discretization  used for the small scene. The blue squares correspond to the each zone, while the numbers correspond to the ID given to each zone.}
	\label{fig:stage_division}
\end{figure} 
%TODO Change this paragraph and motivate the reason this is important and instead of puting the figures, write the references to those figures
The scene's description is the following: the robot starts in the middle of the stage to move to the upstage right (Figure~\ref{fig:StageDirections}), close to the right wing. Then, the robot moves to upstage right center and rotates by $\pi/2$ left (Figure~\ref{fig:BodyPosition}). Next the robot moves to the right center to then go to the center. When it arrives there, it turns full back and move backwards to downstage center with a full front orientation. There, it turns full back to move to center. Finally the robot turns to profile right and it does a step back; then it goes to the upstage center and then upstage right. The sequence of movements programmed to the robot are depicted in Figure~\ref{fig:movement}.
\begin{figure*}
	\centering
	\includegraphics[width=0.95\textwidth]{./Images/fourthCaseSceneD.png} 
	\caption{Sequence of movements done by the robot. The red arrows show the trajectory done by the robot, while the numbers show the order among the movements. a) The first ten movements b) The last five movements }
	\label{fig:movement}
\end{figure*}

>>>>

The case study was done at Researcher's Night 2015 with two main objectives (i) cross-validate the findings obtained from the experiment, and (ii) use the Emotional Enrichment System to verify whether the participants would prefer scenes when the robot expresses emotions or rather moves without any emotion expression.

@@

This case study uses the results obtained in the experiment and was designed to have two parts (i) emotion expression through changes in linear velocity, angular velocity, oscillation angle, orientation and direction.  
(ii) Presentation of a small scene to verify whether the participants would prefer scenes when the robot shows emotions or not. the Emotional Enrichment System was used in both cases (with and without emotions). Two web-cams and eight Alvar tags were added to use Kalman filter to improve robots localization in the stage. The detection of the AR tags is done through the use of the ROS package ar\_track\_alvar~\cite{artag2015}. The distribution of the web-cams and the tags are depicted in Figure~\ref{fig:setup_fourth}.

@@

\begin{table}[h]
\centering
\small
\caption{Parameters' values selected from the experiment.}
		\label{table:selected_fourth}
		\begin{tabular}{|c|p{0.9 cm}|p{0.9 cm}|p{0.9 cm}|p{1.05 cm}|p{0.9 cm}|}
			\hline
%\rotatebox{90}{\textbf{Emotion } }&
%\rotatebox{90}{\textbf{Direction  ($rad$)}}&
%\rotatebox{90}{\textbf{Orientation ($rad$)} }&
%\rotatebox{90}{\textbf{Linear Velocity ($mm/s$) }}&
%\rotatebox{90}{\textbf{Angular Velocity ($rad/s$) }}&
%\rotatebox{90}{\textbf{Angle ($rad$)}}\\	
\textbf{Emotion}&\textbf{Direc-tion  ($rad$)} & \textbf{Orien-tation ($rad$)} & \textbf{Linear Velocity ($mm/s$) } & \textbf{Angular Velocity ($rad/s$) } & \textbf{Angle ($rad$)} \\
			\hline
			Happiness 1&$0$&$0$&$500$&$3$&$0.349$\\
			\hline
			Happiness 2&$0$&$0$&$900$&$3$&$0.174$\\
			\hline
			Anger 1&$\pi$&$0$&$500$&$3$&$0.087$\\
			\hline
			Anger 2&$0$&$0$&$900$&$1$&$0.087$\\
			\hline
			Fear 1&$\pi$&$\pi$&$900$&$2$&$0.174$\\
			\hline
			Fear 2&$\pi$&$\pi$&$500$&$2$&$0.087$\\
			\hline
			Sadness 1&$\pi$&$0$&$200$&$1$&$0.349$\\
			\hline
			Sadness 1&$0$&$\pi$&$200$&$1$&$0.349$\\
			\hline
			\end{tabular}
\end{table}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/FourthCaseScene.png} 
	\caption{Stage discretization  used for the small scene. The blue squares correspond to the each zone, while the numbers correspond to the ID given to each zone.}
	\label{fig:stage_division}
\end{figure} 
%TODO Change this paragraph and motivate the reason this is important and instead of puting the figures, write the references to those figures
The scene's description is the following: the robot starts in the middle of the stage to move to the upstage right (Figure~\ref{fig:StageDirections}), close to the right wing. Then, the robot moves to upstage right center and rotates by $\pi/2$ left (Figure~\ref{fig:BodyPosition}). Next the robot moves to the right center to then go to the center. When it arrives there, it turns full back and move backwards to downstage center with a full front orientation. There, it turns full back to move to center. Finally the robot turns to profile right and it does a step back; then it goes to the upstage center and then upstage right. The sequence of movements programmed to the robot are depicted in Figure~\ref{fig:movement}.
\begin{figure*}
	\centering
	\includegraphics[width=0.85\textwidth]{./Images/fourthCaseSceneD.png} 
	\caption{Sequence of movements done by the robot. The red arrows show the trajectory done by the robot, while the numbers show the order among the movements. a) The first ten movements b) The last five movements }
	\label{fig:movement}
\end{figure*}

### 2016-09-26 HRI/Parts/System.tex
Small correction in the Antecedents, which is a description of the experiment, the system. There is still need work on the case study description.

The system used is divided in the platform and emotional enrichment system.

@@

A non-anthropomorphic robotic platform without any bio-inspired appearance was developed and used in this case study. The holonomic platform was built using Odroid U3, Arduino Due, and 3 metal gear motors with 64 CPR encoders and omniwheels. The platform could be observed in Figure~\ref{fig:Robot}

@@

The main idea of Emotional Enrichment System (EES) is to blend an emotion and a action to produce an "emotional action". However this is not far as previous works have done so far. The system has been envisioned to also allow:

@@

In other words, the system could be thought as a box that receives a desire action, action's parameters and emotions to blend them together without paying particular attention who or how the decision to execute a particular action with specific parameters were made. This is achieved through the use of text messages to describe the action that should be executed, as well its parameters, emotion and emotion's intensity.  To change the emotion without affecting the action that is being executed, these information are divided into two messages: one for the action and its parameters, and other for the emotion and its intensity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Simple and Compound Actions}
To achieve actions' enrichment with emotions, two different types of actions are used: simple and compound actions. Simple actions are actions that could be considered as ''atomic''; these actions can be enriched with emotion changing their specific parameters. For example, consider two simple actions: speak and move body. The human action of speaking is related to changes in the vocal cords to produce different sounds, while move body is directly related to the body movement. Therefore, for the speaking action the following parameters are expected: text, pitch and tone. On the other hand, the action move body would have as parameters the desired destination, velocity, and trajectory constraints. Since each action is characterized by different parameters, the parameters' modifications to obtain an emotional action will be different for each of them.

\subsubsection{Emotional Execution Tree}
The \textit{Emotional Enrichment System}  is grounded on the use of \textit{Emotional Execution Tree} ($EXT$), which is a computational representation of actions desired actions that should be executed. 
$EXT$ is a connected acyclic graph with vertices and  edges. The root and non-leaf nodes could be \textit{parallel} or \textit{sequence} type. The parallel node could be one out of four different sub-types: action and emotion synchronous, action synchronous and emotion asynchronous, action asynchronous and emotion synchronous, or action and emotion asynchronous. Sequential nodes could just be one of two sub-types: emotion synchronous or asynchronous. Action synchronous means that each time that a parallel node receives a finish notification (success or failure), it will send a finish message to all the nodes that derived from it. On the other hand, emotion synchronous means that each time that a node (sequence or parallel) receives an emotion synchronization message, it will propagate the message to all the actions in the branches. 
This distinction creates the possibility to synchronize emotional changes without affecting the normal execution of the desired action. Finally, the leaf nodes could only be simple action nodes. All the nodes can assume two levels: principal or secondary. If a node is principal, it will notify its predecessor about the messages that it has received, while the secondary cannot propagate any message to its predecessor.

@@

The process to enrich an action is threefold as follows:
\begin{enumerate}

	\item \textit{Generation of emotional execution tree:} this phase starts every time that a new action message is received. The process begins by parsing the format, verifying that the actions described on it exist in the system, and that the parameters correspond to the ones expected by each action described on the message. This parameters' verification is done on the implemented description for each action, which describes the parameters that are mandatory and those that are optional. When the verification is done, and all the action exists and  the parameters correspond, an $ext$ is created.

	\item \textit{Emotion addition:} uses the $ext$ created in the previous phase. In this phase new $sa$ are added to the $ext$ and the $sa$'s parameters are modified following the emotion description, which is loaded from files. This process is broken down in two steps. First, all the actions that are required to convey the desired emotion, and that are not yet present are added. Second, the emotional parameters are modulated based on the emotion's intensity and character traits.

>>>>

The same system used in the experiment was used in the case study presented in this paper and they are explained in the following subsection.

@@

The non-anthropomorphic robotic platform was created to do not have any bio-inspired appearance. The holonomic platform was built using Odroid U3, Arduino Due, and 3 metal gear motors with 64 CPR encoders and omniwheels. The platform could be observed in Figure~\ref{fig:Robot}. The Arduino Due is in charge to be the interface between the hardware (e.g. motors) and Odroid, which host all the emotion enrichment system.

@@

The main idea of Emotional Enrichment System (EES) is to blend an emotion and a action to produce an "emotional action". However this is not far as previous works have done so far. Therefore, the system has been envisioned to also allow:

@@

In other words, the system could be thought as a box that receives a desire action, action's parameters and emotion to blend them together without paying particular attention on who or how the decision to execute a the particular action was made. This is achieved through the use of messages to describe the action that should be executed, as well its parameters, emotion and emotion's intensity.  These information are divided into two messages: one for the action and its parameters, and other for the emotion and its intensity. This is done to do not affect the action's execution.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Simple and Compound Actions}

To achieve actions' enrichment with emotions, two different types of actions are used: simple and compound actions. Simple actions are actions that could be considered as ''atomic''; these actions can be enriched with emotion changing their specific parameters. For example, consider two simple actions: speak and move body. The human action of speaking is related to changes in the vocal cords to produce different sounds, while move body is directly related to the body movement. Therefore, for the speaking action the following parameters are expected: text, pitch and tone. On the other hand, the action move body would have as parameters the desired destination, velocity, and trajectory constraints. Since each action is characterized by different parameters, the parameters' modifications to obtain an emotional action will be different for each of them. On the other hand, compound action is an action created from other compound or simple actions. For example, suppose that it is required that the robot has to \textit{recite} some dialogue and \textit{walk} to a different position, as doing a trajectory. A possible implementation could consist on speak simple action in parallel to two consecutive action walk, which is also a compound action that is composed by three parallel actions: balance left arm, balance right arm, and move body.

Although compound actions could be used to generate diversity of actions from simple actions, the main drawback of this approach is that all possible combination of actions should code in the system to be used. Using the example previously described, if it is decided to have three positions instead of two, then it would mean that a new compound action should be implemented to have an action in which could be specified three positions. To overcome this limitation, it was created a language to specify actions that are not implemented in the system. A computational representation of these language is used to execute the actions described in this language.

\subsubsection{Emotional Execution Tree}
The \textit{Emotional Enrichment System}  is grounded on the use of \textit{Emotional Execution Tree} ($EXT$), which is a computational representation of desired actions that must be executed. 
$EXT$ is a connected acyclic graph with vertices and  edges. The root and non-leaf nodes could be \textit{parallel} or \textit{sequence} type. The parallel node could be one out of four different sub-types: action and emotion synchronous, action synchronous and emotion asynchronous, action asynchronous and emotion synchronous, or action and emotion asynchronous. Sequential nodes could just be one of two sub-types: emotion synchronous or asynchronous. Action synchronous means that each time that a parallel node receives a finish notification (success or failure), it will send a finish message to all the nodes that derived from it. On the other hand, emotion synchronous means that each time that a node (sequence or parallel) receives an emotion synchronization message, it will propagate the message to all the actions in the branches. 
This distinction creates the possibility to synchronize emotional changes without affecting the normal execution of the desired action. Finally, the leaf nodes could only be simple action nodes. All the nodes can assume two levels: principal or secondary. If a node is principal, it will notify its predecessor about the messages that it has received, while the secondary cannot propagate any message to its predecessor.

@@

The action enrichment is achieved through the following phases:
\begin{enumerate}

	\item \textit{Generation of emotional execution tree:} this phase starts every time that a new action message is received. The process begins by parsing the format, verifying that the actions described on it exist in the system, and that the parameters correspond to the ones expected by each action described in the message. This parameters' verification is done on the implemented description for each simple action, which describes the parameters that are mandatory and those that are optional. When the verification is done, and all the action exists and  the parameters correspond, an $ext$ is created.

	\item \textit{Emotion addition:} uses the $ext$ created in the previous phase. In this phase new simple actions are added to the $ext$ and the simple action's parameters are modified following the emotion description, which is loaded from files. This process is broken down in two steps. First, all actions that are required to convey the desired emotion, and that are not yet present in the branch are added. Second, the emotional parameters are modulated based on the emotion's intensity and character traits.
